Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-01-21 ms.tex
more clean up

To incorporate measurements on the Wien side of the dust emission spectrum, 
we include 100$\mu$m data in our SED fits. In particular, we use the SFD 
reprocessing of \IRAS~100$\mu$m, which we will refer to as \verb|i100|, and at 
times by frequency as 3000 GHz. The \verb|i100| map has angular resolution of 
$6.1'$, and was constructed so as to contain only thermal 
emission from Galactic dust, with compact sources and zodiacal light removed, 
and its zero level tied to HI. We use the \verb|i100| map as-is, without any 
custom modifications.
%maybe clarify that throughout we use SFD i100 totally as-is, with no
% customization

@@

%maybe one sentence here stating that the goal of next few subsections
%is to isolate Galactic thermal dust emission

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}
In preparing the \PLANCK~maps for thermal dust SED fitting, we first addressed
the CMB anisotropies before performing any of the interpolation/smoothing 
described in $\S$\ref{sec:ptsrc}/$\S$\ref{sec:smth}. The CMB anisotropies are 
effectively imperceptible in \PLANCK~857 GHz, but can be perceived at a low 
level in \PLANCK~545 GHz, and are prominent in \PLANCK~217 GHz and 353 GHz 
relative to the Galactic emission we wish to characterize, especially at high 
latitudes. To remove the CMB anisotropies, we have subtracted the Spectral 
Matching Independent Component Analysis (\verb|SMICA|) model from each of the 
\PLANCK~maps \citep{smica}, after applying appropriate unit conversions. 
Removal of the residual Solar dipole at low frequencies is discussed in 
$\S$\ref{sec:zp}.
% SMICA is subtracted from ALL Planck maps being used, including 545, 857 GHz

@@

\subsection{Zero-Point}
\label{sec:zp}
Although we wish to isolate and model thermal emission from Galactic dust, the
\PLANCK~maps contain additional components on large angular scales. At each 
frequency, there can exist an overall, constant offset that must be subtracted 
to set the zero level of Galactic dust. Additionally, faint residuals of the 
Solar dipole remain at low frequencies. At 353 GHz, 545 GHz, 857 GHz, we have 
adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
 by correlating \PLANCK~857 GHz versus HI at low column density, and then by 
correlating the other \PLANCK~channels against the corrected 857 GHz map. We 
have not altered the \verb|i100| zero-point in any way. We have also removed 
the Solar dipole residual according to \cite{planckdust}, by subtracting a 
dipole pattern oriented toward $(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and 
scaled according to their Table 1.

@@

The offset and residual dipole parameters for \PLANCK~217 GHz were not 
available from \cite{planckdust} Table 1. In order to set the zero-point of 
Galactic emission in the 217 GHz map, we simultaneously fit the constant 
offset and residual dipole amplitude. Restricting to regions of low column 
density (HI $<$ 200 K km/s), we modeled the 217 GHz map as:

@@

With $i$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low HI cut, we have an overconstrained linear model with three
parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improvement striking in this scatterplot, but the 
residual dipole also appears visually well-corrected in the 217 GHz map 
itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

@@

\subsection{Finite Bandpass Correction}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. But in reality
we are comparing to broad bands typically hundreds of GHz wide. The shape of 
model spectrum $M_{\nu}$ is only dependent upon one parameter: the hot
dust temperature $T_2$. We have therefore computed bandpass correction factors
$b_{\nu_i}$ such that $M_{\nu_i}b_{\nu_i}$ is the predicted observation for the
\PLANCK~or \IRAS~band of interest. We have used the \PLANCK~transmission 
curves provided in the file \verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. 
For 3000 GHz, we have used the \IRAS~bandpass, following the treatment of
\cite{planckdust}. We note that it is not clear which bandpass should
properly be considered for the \IRAS/DIRBE composite 100$\mu$m maps used in 
this analysis (here SFD \verb|i100|) and that of \cite{planckdust}.

>>>>

To incorporate measurements on the Wien side of the dust emission spectrum, 
we include 100$\mu$m data in our SED fits. In particular, we use the SFD 
reprocessing of \IRAS~100$\mu$m, which we will refer to as \verb|i100|, and at 
times by frequency as 3000 GHz. The \verb|i100| map has angular resolution of 
$6.1'$, and was constructed so as to contain only thermal 
emission from Galactic dust, with compact sources and zodiacal light removed, 
and its zero level tied to H\,\textsc{i}. We use the \verb|i100| map as-is, without any 
custom modifications.
%maybe clarify that throughout we use SFD i100 totally as-is, with no
% customization

@@

The following subsections detail the processing steps we have applied to 
isolate thermal emission from Galactic dust in the \PLANCK~maps and prepare
these maps for our joint \IRAS/\PLANCK~SED fits.

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}
We first addressed the CMB anisotropies before performing any of the 
interpolation/smoothing described in $\S$\ref{sec:ptsrc}/$\S$\ref{sec:smth}. 
The CMB anisotropies are effectively imperceptible in \PLANCK~857 GHz, but can 
be perceived at a low level in \PLANCK~545 GHz, and are prominent in 
\PLANCK~217 GHz and 353 GHz relative to the Galactic emission we wish to 
characterize, especially at high latitudes. To remove the CMB anisotropies, we 
have subtracted the Spectral Matching Independent Component Analysis 
(\verb|SMICA|) model from each of the \PLANCK~maps \citep{smica}, after 
applying appropriate unit conversions. Removal of the residual Solar dipole at 
low frequencies is discussed in $\S$\ref{sec:zp}.
% SMICA is subtracted from ALL Planck maps being used, including 545, 857 GHz

@@

\subsection{Zero-Point}
\label{sec:zp}
Although we wish to isolate and model thermal emission from Galactic dust, the
\PLANCK~maps contain additional components on large angular scales. At each 
frequency, there can exist an overall, constant offset that must be subtracted 
to set the zero level of Galactic dust. Additionally, faint residuals of the 
Solar dipole remain at low frequencies. At 353 GHz, 545 GHz, 857 GHz, we have 
adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
 by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
then by correlating the other \PLANCK~channels against the corrected 857 GHz 
map. We have not altered the \verb|i100| zero-point in any way. We have also 
removed the Solar dipole residual according to \cite{planckdust}, by 
subtracting a dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
Table 1.

@@

The offset and residual dipole parameters for \PLANCK~217 GHz were not 
available from \cite{planckdust} Table 1. In order to set the zero-point of 
Galactic emission in the 217 GHz map, we simultaneously fit the constant 
offset and residual dipole amplitude. Restricting to regions of low column 
density (H\,\textsc{i} $<$ 200 K km/s), we modeled the 217 GHz map as:

%in previous sentence, mention the specific cut on HI velocities considered

@@

With $i$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low H\,\textsc{i} cut, we have an overconstrained linear model with
three parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improved correlation striking in this scatterplot, 
but the residual dipole also appears visually well-corrected in the 217 GHz 
map itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

@@

\subsection{Finite Bandpass Correction}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. But in reality
we are comparing to broad bands typically hundreds of GHz wide. The shape of 
model spectrum $M_{\nu}$ is only dependent upon one parameter: the hot
dust temperature $T_2$. We have therefore computed bandpass correction factors
$b_{\nu_i}$ such that $M_{\nu_i}b_{\nu_i}$ is the predicted observation for the
\PLANCK~or \IRAS~band of interest. We have used the \PLANCK~transmission 
curves provided in the file \verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. 
For 3000 GHz, we have used the \IRAS~bandpass, following the treatment of
\cite{planckdust}. We note that it is not clear which bandpass should
properly be ascribed to the \IRAS/DIRBE composite 100$\mu$m maps under
consideration. % (here SFD \verb|i100|) and that of \cite{planckdust}).
